Drought indices of water demand and supply, soil moisture, vegetation, and their impact on LULCC in continental Chile

### 2024-01-23 article/paper_drought_Chile2023.qmd
actualizar repo gema
actualizar repo depto

###	in-situ data

POR COMPLETAR

@@

POR COMPLETAR

@@

@@

## Ecological and agricultural drought

@@

2.- Sobre los tipos de land cover más afectados por los indicadores de sequía. Asociación con el matorral chileno [@Fuentes2021]. Diferencia entre el Norte Chico, Centro y lo que pasa hacía el sur.

>>>>

###	Weather stations

We validated the ERA5L variables for monthly mean temperature, total precipitation, and volumetric soil water content. For temperature and precipitation, we used the weather network from the Ministry of Agriculture of Chile (www.agromet.com). We used 266 stations located throughout Chile. For soil moisture, we select a soil network for water content and temperature, localized in the Aconcagua's river basin ("Norte Chico"). This network corresponds to 53 sensors of the type Temperature Moisture Sensor (TMS) [@Wild2019].

@@

To account for the performance of the ERA5L climatic variables regarding the values measured by the weather stations. We selected the following metrics:

$$MAE = \frac{1}{n}\sum |{E-S}|$$
$$Bias = \frac{\sum E}{\sum S}$$
$$ubRMSE =\sqrt{\frac{\sum{ \left[ (E_i-\overline{E})-(S_i-\overline{S}) \right ] ^2}}{n}}$$

$$CC = \frac{\sum (S_i-\bar{S})(E_i-\bar{E})}{\sqrt{(Si-\bar{S})^2(E_i-\bar{E})^2}}$$

$MAE$: mean absolute error
$bias$: bias
$ubRMSE$: unbiassed root mean squared error
$CC$: coefficient of correlation
$S$: value of the variable measure by the weather station
$E$: value of the variable measure by ERA5L

@@

The average metrics of performance of ERA5L over the 266 weather stations were in the case of temperature: $ubRMSE=1.06\,°C$, $MAE=1.131\,°C$, and $CC=0.963$. For precipitation, $MAE=28.1\,mm$, $Bias=1.93$, and $CC=0.845$.

@@

@@

2.- Sobre los tipos de land cover más afectados por los indicadores de sequía. Asociación con el matorral chileno [@Fuentes2021]. Diferencia entre el Norte Chico, Centro y lo que pasa hacía el sur (@Miranda2023)
